# Results
```{r}
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
suppressMessages(library(knitr))
suppressMessages(suppressWarnings(library(janitor)))
suppressMessages(library(scales))
BroadwayData <- read_csv("Raw_Data/broadway.csv", show_col_types = FALSE)
colnames(BroadwayData) <- c("Day", "Date", "Month","Year","Show_Name","Show_Theatre","Show_Type","Attendance","Capacity","Gross","Gross_Potential","Performances")
```
1) Linear graph of Attendance vs Gross or gross potential (Cooper)

2) Do the above, but faceted by show.type (Cooper) 

3) Gross by Month and/or year (Cooper) - Test out combinations (facet by year?)

4) facet again by show.type (Cooper) 

5) Attendance by year and/or month (Aylmer) 
```{r}
attendance_by_year <- BroadwayData |> 
  group_by(Year) |> 
  summarize(total_attendance = sum(Attendance), .groups = "drop")

ggplot(attendance_by_year, aes(x = Year, y = total_attendance)) +
  geom_col(fill = "skyblue") +
  labs(title = "Total Attendance by Year", x = "Year", y = "Total Attendance Count")+
  scale_y_continuous(labels = label_number(big.mark = ","),
                     breaks = seq(0, max(attendance_by_year$total_attendance), by = 2500000))
```

```{r}
attendance_by_month <- BroadwayData |> 
  group_by(Month) |> 
  summarize(avg_attendance = mean(Attendance), .groups = "drop")

attendance_by_month$Month <- month.abb[attendance_by_month$Month]
attendance_by_month$Month <- factor(attendance_by_month$Month, levels = month.abb)

ggplot(attendance_by_month, aes(x = Month, y = avg_attendance)) +
  geom_col(fill = "skyblue") +
  labs(title = "Average Attendance by Month", x = "Month", y = "Average Attendance Count")+
  scale_y_continuous(labels = label_number(big.mark = ","),
                     breaks = seq(6000, 10000, by = 500))+
  coord_cartesian(ylim = c(7000, 9000))
```
6) swap 3 and 4's faceting of show type and month (i.e facet by month, plot the 3 show types) (Cooper) 

7) Stacked bar chart attendance & capacity vs year Exclude (Aylmer)

8) Stacked bar chart attendance & capacity for each theatre (Aylmer) 

9) Interactive idea for now: 1) show some trend year by year (animation or selection tool) or 2) Put together an interactive predictor