# Results
```{r}
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
suppressMessages(library(knitr))
suppressMessages(suppressWarnings(library(janitor)))
BroadwayData <- read_csv("Raw_Data/broadway.csv", show_col_types = FALSE)
colnames(BroadwayData) <- c("Day", "Date", "Month","Year","Show_Name","Show_Theatre","Show_Type","Attendance","Capacity","Gross","Gross_Potential","Performances")
```
1) Linear graph of Attendance vs Gross or gross potential (Cooper)

2) Do the above, but faceted by show.type (Cooper) 

3) Gross by Month and/or year (Cooper) - Test out combinations (facet by year?)

4) facet again by show.type (Cooper) 

5) Attendance by year and/or month (Aylmer) 
```{r}
ggplot(BroadwayData, aes(A))
```
6) swap 3 and 4's faceting of show type and month (i.e facet by month, plot the 3 show types) (Cooper) 

7) Stacked bar chart attendance & capacity vs year (Aylmer)

8) Stacked bar chart attendance & capacity for each theatre (Aylmer) 

9) Interactive idea for now: 1) show some trend year by year (animation or selection tool) or 2) Put together an interactive predictor